Comoving stars in \textsl{Gaia DR1}:\\ An abundance of very wide separation comoving pairs

### 2017-06-01 paper/ms.tex
put in new columns and make schema standalone
add flatiron
remove sentence
minor punctuations edit from the proof

\title{Co-moving stars in \textsl{Gaia DR1}:\\
  An abundance of very wide separation co-moving pairs}
\author{Semyeong Oh\altaffilmark{\pu,\lead},
        Adrian M. Price-Whelan\altaffilmark{\pu},
        David W. Hogg\altaffilmark{\ccpp,\mpia,\cca},
        Timothy D. Morton\altaffilmark{\pu},
        David N. Spergel\altaffilmark{\pu,\cca}
}

@@

\altaffiltext{\pu}{Department of Astrophysical Sciences,
                   Princeton University, Princeton, NJ 08544, USA}
\altaffiltext{\lead}{To whom correspondence should be addressed:
                     \texttt{semyeong@astro.princeton.edu}}
\altaffiltext{\ccpp}{Center for Cosmology and Particle Physics,
                     Department of Physics,
                     New York University, 4 Washington Place,
                     New York, NY 10003, USA}
\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
                     K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{\cca}{Center for Computational Astrophysics, 162 5th Ave, New York, NY 10003, USA}

\begin{abstract}
The primary sample of the \gaia\ Data Release 1 is the
\textsl{Tycho-Gaia Astrometric Solution} (\tgas): $\approx$ 2 million Tycho-2
sources with improved parallaxes and proper motions relative to the initial
catalog.
This increased astrometric precision presents an opportunity to find new binary
stars and moving groups.
We search for high-confidence co-moving pairs of stars in \tgas\ by identifying
pairs of stars consistent with having the same 3D velocity using a marginalized
likelihood ratio test to discriminate candidate co-moving pairs from the field
population.
Although we perform some visualizations using (bias-corrected) inverse-parallax
as a point-estimate of distance, the likelihood ratio is computed with a
probabilistic model that includes the covariances of parallax and proper motions,
and marginalizes the (unknown) true distances and 3D velocities of the stars.
We find 13,085 co-moving star pairs among 10,606 unique stars with separations
as large as 10 pc (our search limit).
Some of these pairs form larger groups through mutual co-moving neighbors:
many of these pair networks correspond to known open clusters and OB
associations, but we also report the discovery of several new co-moving groups.
Most surprisingly, we find a large number of very wide ($>1$ pc) separation
co-moving star pairs, the number of which increases with increasing separation
and cannot be explained purely by false-positive contamination.
Our key result is a catalog of high-confidence co-moving pairs of stars in
\tgas.
We discuss the utility of this catalog for making dynamical inferences about the
Galaxy, testing stellar atmosphere models, and validating chemical abundance
measurements.
\end{abstract}

@@

Stars that are roughly co-located and moving with similar space velocities
(``co-moving stars'') are of special interest in many branches of astrophysics.

@@

Wide binaries are also good testbeds for stellar models and age indicators:
the constituent stars were likely born at the same time with the same chemical
compositions, but evolved independently because of their wide separation.
These pairs are therefore useful for validating gyrochronology relations
(e.g., \citealt{Chaname:2012}) and may be valuable for testing consistency
between stellar atmosphere models.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low mass star and a
much brighter F/G/K star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

At larger separations ($\gtrsim 1$~pc), co-moving stars are likely members of
(potentially dissolving) moving groups, associations, and star clusters or
disrupted wide binaries.
The origin of moving groups is still under active debate (e.g., \citealt{Bovy:2010aa}):
are they remnants of a coeval
star formation event with similar chemical composition? Or are they formed by
dynamical effects of non-axisymmetric features of the Galaxy such as spirals
and bars? With the recent advances in measuring chemical abundances of
a large volume of stars using high and low resolution spectroscopy
(e.g., \citealt{Steinmetz:2006aa,Majewski:2015aa,Gilmore:2012aa} to name a few),
we can now start to explore these questions
with unprecedented statistics and in unexplored detail.
The dynamics of cluster dissolution provides important clues to understanding
the star formation history and the dynamical evolution of the Milky Way.
In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
(``stellar streams''; see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
These tidal streams are modeled to infer the parameters of the Galactic
potential (e.g., \citealt{Kupper:2015}).
Similar processes are at work with star clusters in the disk.
However, the dynamical time is much shorter, and the dynamics will be much more
complex because of the existence of other perturbers in the disk.

To date, thousands of candidate co-moving star pairs have been identified by
searching for stars with common proper motions
(\citealt{Poveda:1994aa,Allen:2000aa,Chaname:2004aa,
Lepine:2007aa,Shaya:2011aa,Alonso-Floriano:2015aa}).
Here, we use the recent first data release of \gaia\ which includes precise distances,
enabling us to ask whether two stars share the same \emph{physical (3D) velocity} rather than
just the projections in the proper motion space.

This paper proceeds as follows:
In \sectionname~\ref{sec:data}, we briefly describe the data set used in
this work.
In \sectionname~\ref{sec:methods}, we develop a statistical method to identify
high-confidence co-moving pairs in this catalog.
In \sectionname~\ref{sec:results}, we present and discuss our resulting catalog
of co-moving pairs.
We summarize in \sectionname~\ref{sec:summary}.

@@

We construct an initial sample of star pairs to search for co-moving pairs as
follows.
We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to the \tgas,
which leaves 619,618 stars.
Then, for each star we establish an initial sample of possible
co-moving partners by selecting all other stars that satisfy two criteria:
separation less than 10~pc and difference
in (point-estimate) tangential velocity less than $\absdvtan < 10$~\kms.
We ultimately build a statistical model that incorporates the covariances of the
data, but for these initial cuts and for visualizations we use a point-estimate
of the distance by applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  \hat{r} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
    \label{eq:dist}
\end{equation}
where $\varpi$ is the parallax in mas.
An estimate for the difference in tangential velocity between two stars is,
then,
\begin{equation}
  \absdvtan = |\hat{r}_1 \vec\mu_1 - \hat{r}_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_{\alpha^*}, \mu_\delta)$.
\footnote{$\mu_{\alpha^*}$ is the proper motion component
in the right ascension direction, $\mu_{\alpha^*} = \mu_\alpha \cos\delta$}

@@

\begin{itemize}
  \item At small separations ($<1$~pc), there is a population of pairs with
    very small tangential velocity difference ($<2$~km/s). Given that these
    pairs are very close in both 3D position and tangential velocities, it is highly
    probable that they are actually co-moving wide binaries.

  \item  A sample of co-moving stars also include stars that are part of, e.g.,
    OB associations, moving groups, and open clusters.
    These astrophysical objects may be detected as a network of co-moving pairs,
    sharing some mutual co-moving neighbors.
    As the pair separation increases, the nature of co-moving pairs
    will change from binaries to those related to these larger objects,
    which generally subtend a larger angle in sky.
    Since the proper motions of two stars with the same 3D velocity
    are projections of this velocity onto the celestial sphere at
    two different viewing angles,
    the larger the difference in viewing angles is, the larger the difference in tangential
    velocities will be.
    Due to this projection effect, a population of genuine co-moving pairs
    will extend to larger \absdvtan\ at larger separation.
    This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
    in the lower right corner that gets thinner as \absdvtan\ increases.

  \item Finally, there is a population of ``random'' pairs of field stars
    that are not co-moving, but still have $\absdvtan < 10$~\kms\
    by chance.
    As \absdvtan\ increases, this population will dominate.
    Figure~\ref{fig:dv-sep} shows that there is an overlap between
    genuine co-moving pairs and ``random'' pairs.
\end{itemize}

In the following section, we construct a statistical model that propagates
the non-trivial uncertainties in the data to our beliefs about the likelihood
that a given pair of stars is co-moving.

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving pairs in the \tgas\ data at a range
of separations.
Here, we develop a method to select high-confidence co-moving
pairs that properly incorporates the uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = \transp{(\mu_{\alpha^*}\, \mu_\delta)}$
    ---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3D velocities of stars in a given pair
    $(\vec{v}_i, \vec{v}_j)$ (relative to the solar system
    barycenter) are either (1) the same with a small (Gaussian) dispersion $s$
    or (2) independent.
    In both cases, velocity is drawn from
    the velocity prior $p(\vec{v})$.
\end{itemize}

@@

For a given pair, we compute the fully marginalized likelihood (FML) for the hypotheses (1) and (2),
$\mathcal{L}_1$ and $\mathcal{L}_2$.
We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as the scalar quantity to select
candidate co-moving pairs, as described in $\sectionname~\ref{sub:selection}$ in
more detail.
To compute these FMLs, the likelihood functions for each star in a pair, $L_i,
L_j$, are marginalized over the (unknown) true distance and 3D velocity for each star in
the pair $(i,j)$.
\begin{align}
  \mathcal{L}_1 &=
    \int \, \dd r_i \, \dd r_j \, \dd^3 \vec{v} \,
    L_i(\vec{\mu_i} \given \vec{v}, r_i, s^2) \,
    L_j(\vec{\mu_j} \given \vec{v}, r_j, s^2) \,
    p(\vec{v}) \, p(r_i \given \varpi_i) \, p(r_j \given \varpi_j) \label{eq:hyp1}\\
  \mathcal{L}_2 &=
    \int \, \dd r_i \, \dd r_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
    L_i(\vec{\mu_i} \given \vec{v}_i, r_i, s^2) \,
    L_j(\vec{\mu_j} \given \vec{v}_j, r_j, s^2) \,
    p(\vec{v}_i) \, p(\vec{v}_j) \, p(r_i \given \varpi_i) \, p(r_j \given \varpi_j). \label{eq:hyp2}
\end{align}
Here, $p(r_i \given \varpi_i)$ is the posterior distribution of distance given parallax measurement
$\varpi_i$ and its Gaussian error $\sigma_{\varpi,i}$.
Note that the FML for the hypothesis (1) involves integration over one velocity $\vec{v}$ that
generates the likelihoods for both stars, $L_i$ and $L_j$.
The marginalization integral for hypothesis (2) can be split into the product of
two simpler integrals $\mathcal{L}_2 = Q(\vec{\mu_i}, \varpi_i) \, Q(\vec{\mu_j}, \varpi_j)$ where
\begin{equation}
  Q(\vec{\mu}, \varpi) = \int \, \dd r \, \dd^3 \vec{v} \, L(\vec{\mu} \given \vec{v}, r, s^2) \,
  p(\vec{v}) \, p(r \given \varpi)
\end{equation}

@@

This section is divided into three parts. First, we discuss and justify a cut of
the likelihood ratio to select candidate co-moving pairs.
Second, we present the statistics and properties of our candidate co-moving
pairs.
Finally, we describe our catalog of candidate co-moving pairs, the main product
of this study.

\subsection{Selecting candidate co-moving pairs}
\label{sub:selection}

In this section, we examine the distribution of (log-)likelihood ratios
$\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
for this quantity to select co-moving pairs from the initial sample.

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{likelihoodratios.pdf}
  \end{center}
  \caption{%
    Justification of the likelihood ratio cut.
    On the top row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
    changes (left) in slices of \absdvtan\ (sequentially color-coded in pink on the right panel).
    On the bottom row, we show how the distribution of pairs on \absdvtan\ vs. separation changes (right)
    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue on the left).
    The black dots in the upper right panel and
    the black line in the lower left panel correspond to the entire initial sample of pairs.
    For comparison, we also present the likelihood ratio distribution of random pairs
    of stars in the \tgas\ with parallax $S/N>8$ in gray filled histogram.
    Based on this, we choose $\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as our high-confidence
    candidate co-moving pairs.
    \label{fig:likelihoodratios}}
    \vspace{1em}
\end{figure*}

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
$\sim 271$k pairs in the initial sample.
As discussed in \sectionname~\ref{sec:data}, we expect a correlation between the likelihood
ratios of pairs, and their distribution on \absdvtan\ vs separation plane.
Specifically, as we sweep through from small \absdvtan\ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of the likelihood ratios of pairs
in slices of \absdvtan\ (top row of Figure~\ref{fig:likelihoodratios}).
Pairs with $\absdvtan \in (0,2.5)~\kms$ are most likely actual co-moving pairs, and
their likelihood ratio distribution is narrowly peaked at $\gtrsim 5$
(darkest pink histogram in the upper left panel of Figure~\ref{fig:likelihoodratios}).
The distribution peaks at lower values and gets broader as \absdvtan\ increases,
and the number of random pairs increasingly dominate.
On the bottom row of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on \absdvtan\ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \sectionname~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as the initial sample ($S/N > 8$).
Shown as the gray filled histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.

Based on these comparisons, we select candidate co-moving pairs with
$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$.
Out of 271,232 pairs in the initial sample, 13,058 pairs (4.8\%)
satisfy this condition.

\subsection{Statistics and properties of the identified co-moving pairs}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{graphviz_examples.pdf}
  \end{center}
  \caption{%
    Visualizations of a few example connected components of co-moving pairs of stars.
    Each star (node) is marked as a red circle, and a line (edge) is drawn
    between two stars if they are co-moving by our selection criteria
    (see \sectionname~\ref{sub:selection}). On left, we show the largest network
    found in this study corresponding to the Pleiades star cluster.
    On right, we show four examples of connected components with varying sizes.
    The connected component on the upper left panel with a size of 58 corresponds
    to NGC~2632, also known as the Beehive cluster.
    \label{fig:graphviz_examples}}
\end{figure*}

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{glon_d_pie.pdf}
  \end{center}
  \caption{%
    Panoramic view of co-moving pairs of stars around the Sun. This is a cylindrical
    projection of onto the Galactic plane with the Sun at the origin.
    Pairs in connected components of sizes less than 5 are connected by gray lines.
    Pair in connected components of size $\geq 5$ are plotted with a unique random color,
    and their nodes (stars) are highlighted with small black circles.
    We also show the positions of known Milky Way star clusters
    from \citet{Kharchenko:2016aa} as light blue circles of $10$~pc radius,
    and stars in OB associations from \citet{de-Zeeuw:1999aa} as
    blue crosses.
    Some of the larger connected components are clearly associated with known
    clusters, but we also discover quite a few new co-moving groups.
    \label{fig:glon_d_pairlines}}
\end{figure*}

Once we have identified candidate co-moving pairs from the initial
sample, these pairs form an undirected graph where stars are nodes, and edges
between the nodes exist for co-moving pairs of stars.
A star may have multiple co-moving neighbors, and two stars may either
be directly or indirectly connected by a sequence of edges (``path'').
We divide the graph into connected components\footnote{A connected component of an undirected graph $G$
is a subgraph of $G$ in which any two nodes are connected to each other by a path.},
and show the distribution of their sizes in Figure~\ref{fig:hist_ccsize}.
The most common are connected components of size 2, which mean mutually exclusive co-moving
pairs. However, it is clear that there are many aggregates of co-moving stars
discovered by looking for co-moving pairs.
These aggregates are likely moving groups, OB associations, or star clusters.
In 13,058 co-moving pairs that we identified,
there are 4,555 connected components among 10,606 unique stars.
The maximum size of the connected components is 151.
We show this largest connected component along with four other
examples of varying sizes in Figure~\ref{fig:graphviz_examples}.
The largest connected component corresponds to the Pleiades open cluster (left panel of Figure~\ref{fig:graphviz_examples})
while the upper left panel of the right column of Figure~\ref{fig:graphviz_examples}
is NGC~2632, another known Milky Way open cluster \citep{Kharchenko:2016aa}.

>>>>

\title{Comoving stars in \textsl{Gaia DR1}:\\
  An abundance of very wide separation comoving pairs}
\author{Semyeong Oh\altaffilmark{\pu,\lead},
        Adrian M. Price-Whelan\altaffilmark{\pu},
        David W. Hogg\altaffilmark{\ccpp,\mpia,\cca},
        Timothy D. Morton\altaffilmark{\pu},
        David N. Spergel\altaffilmark{\pu,\cca}
}

@@

\altaffiltext{\pu}{Department of Astrophysical Sciences,
                   Princeton University, Princeton, NJ 08544, USA}
\altaffiltext{\lead}{To whom correspondence should be addressed:
                     \texttt{semyeong@astro.princeton.edu}}
\altaffiltext{\ccpp}{Center for Cosmology and Particle Physics,
                     Department of Physics,
                     New York University, 4 Washington Place,
                     New York, NY 10003, USA}
\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
                     K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{\cca}{Center for Computational Astrophysics, Flatiron Institute,
                    162 5th Ave, New York, NY 10003, USA}

\begin{abstract}
The primary sample of the \gaia\ Data Release 1 is the
\textsl{Tycho-Gaia Astrometric Solution} (\tgas): $\approx$ 2 million Tycho-2
sources with improved parallaxes and proper motions relative to the initial
catalog.
This increased astrometric precision presents an opportunity to find new binary
stars and moving groups.
We search for high-confidence comoving pairs of stars in \tgas\ by identifying
pairs of stars consistent with having the same 3D velocity using a marginalized
likelihood ratio test to discriminate candidate comoving pairs from the field
population.
Although we perform some visualizations using (bias-corrected) inverse parallax
as a point estimate of distance, the likelihood ratio is computed with a
probabilistic model that includes the covariances of parallax and proper motions
and marginalizes the (unknown) true distances and 3D velocities of the stars.
We find 13,085 comoving star pairs among 10,606 unique stars with separations
as large as 10 pc (our search limit).
Some of these pairs form larger groups through mutual comoving neighbors:
many of these pair networks correspond to known open clusters and OB
associations, but we also report the discovery of several new comoving groups.
Most surprisingly, we find a large number of very wide ($>1$ pc) separation
comoving star pairs, the number of which increases with increasing separation
and cannot be explained purely by false-positive contamination.
Our key result is a catalog of high-confidence comoving pairs of stars in
\tgas.
We discuss the utility of this catalog for making dynamical inferences about the
Galaxy, testing stellar atmosphere models, and validating chemical abundance
measurements.
\end{abstract}

@@

Stars that are roughly co-located and moving with similar space velocities
(``comoving stars'') are of special interest in many branches of astrophysics.

@@

Wide binaries are also good test beds for stellar models and age indicators:
the constituent stars were likely born at the same time with the same chemical
compositions, but evolved independently because of their wide separation.
These pairs are therefore useful for validating gyrochronology relations
(e.g., \citealt{Chaname:2012}) and may be valuable for testing consistency
between stellar atmosphere models.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low-mass star and a
much brighter F/G/K star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

At larger separations ($\gtrsim 1$~pc), comoving stars are likely members of
(potentially dissolving) moving groups, associations, and star clusters or
disrupted wide binaries.
The origin of moving groups is still under active debate (e.g., \citealt{Bovy:2010aa}):
are they remnants of a coeval
star formation event with similar chemical composition? Or are they formed by
dynamical effects of nonaxisymmetric features of the Galaxy such as spirals
and bars? With the recent advances in measuring chemical abundances of
a large volume of stars using high- and low- resolution spectroscopy
(e.g., \citealt{Steinmetz:2006aa,Majewski:2015aa,Gilmore:2012aa} to name a few),
we can now start to explore these questions
with unprecedented statistics and in unexplored detail.
The dynamics of cluster dissolution provides important clues to understanding
the star formation history and the dynamical evolution of the Milky Way.
In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
(``stellar streams''; see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
These tidal streams are modeled to infer the parameters of the Galactic
potential (e.g., \citealt{Kupper:2015}).
Similar processes are at work with star clusters in the disk.
However, the dynamical time is much shorter, and the dynamics will be much more
complex because of the existence of other perturbers in the disk.

To date, thousands of candidate comoving star pairs have been identified by
searching for stars with common proper motions
(\citealt{Poveda:1994aa,Allen:2000aa,Chaname:2004aa,
Lepine:2007aa,Shaya:2011aa,Alonso-Floriano:2015aa}).
Here, we use the recent first data release of \gaia\ which includes precise distances,
enabling us to ask whether two stars share the same \emph{physical (3D) velocity} rather than
just the projections in the proper motion space.

This paper proceeds as follows:
In \sectionname~\ref{sec:data}, we briefly describe the data set used in
this work.
In \sectionname~\ref{sec:methods}, we develop a statistical method to identify
high-confidence comoving pairs in this catalog.
In \sectionname~\ref{sec:results}, we present and discuss our resulting catalog
of comoving pairs.
We summarize in \sectionname~\ref{sec:summary}.

@@

We construct an initial sample of star pairs to search for comoving pairs as
follows.
We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to the \tgas,
which leaves 619,618 stars.
Then, for each star we establish an initial sample of possible
comoving partners by selecting all other stars that satisfy two criteria:
separation less than 10~pc and difference
in (point-estimate) tangential velocity less than $\absdvtan < 10$~\kms.
We ultimately build a statistical model that incorporates the covariances of the
data, but for these initial cuts and for visualizations we use a point-estimate
of the distance by applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  \hat{r} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
    \label{eq:dist}
\end{equation}
where $\varpi$ is the parallax in mas.
An estimate for the difference in tangential velocity between two stars is,
then,
\begin{equation}
  \absdvtan = |\hat{r}_1 \vec\mu_1 - \hat{r}_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_{\alpha^*}, \mu_\delta)$.
\footnote{$\mu_{\alpha^*}$ is the proper motion component
in the right ascension direction, $\mu_{\alpha^*} = \mu_\alpha \cos\delta$}

@@

\begin{itemize}
  \item At small separations ($<1$~pc), there is a population of pairs with
    very small tangential velocity difference ($<2$~km/s). Given that these
    pairs are very close in both 3D position and tangential velocities, it is highly
    probable that they are actually comoving wide binaries.

  \item  A sample of comoving stars also include stars that are part of, e.g.,
    OB associations, moving groups, and open clusters.
    These astrophysical objects may be detected as a network of comoving pairs,
    sharing some mutual comoving neighbors.
    As the pair separation increases, the nature of comoving pairs
    will change from binaries to those related to these larger objects,
    which generally subtend a larger angle in sky.
    Since the proper motions of two stars with the same 3D velocity
    are projections of this velocity onto the celestial sphere at
    two different viewing angles,
    the larger the difference in viewing angles is, the larger the difference in tangential
    velocities will be.
    Due to this projection effect, a population of genuine comoving pairs
    will extend to larger \absdvtan\ at larger separation.
    This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
    in the lower right corner that gets thinner as \absdvtan\ increases.

  \item Finally, there is a population of ``random'' pairs of field stars
    that are not comoving, but still have $\absdvtan < 10$~\kms\
    by chance.
    As \absdvtan\ increases, this population will dominate.
    Figure~\ref{fig:dv-sep} shows that there is an overlap between
    genuine comoving pairs and ``random'' pairs.
\end{itemize}

In the following section, we construct a statistical model that propagates
the non-trivial uncertainties in the data to our beliefs about the likelihood
that a given pair of stars is comoving.

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of comoving pairs in the \tgas\ data at a range
of separations.
Here, we develop a method to select high-confidence comoving
pairs that properly incorporates the uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = \transp{(\mu_{\alpha^*}\, \mu_\delta)}$
    ---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3D velocities of stars in a given pair
    $(\vec{v}_i, \vec{v}_j)$ (relative to the solar system
    barycenter) are either (1) the same with a small (Gaussian) dispersion $s$
    or (2) independent.
    In both cases, velocity is drawn from
    the velocity prior $p(\vec{v})$.
\end{itemize}

@@

For a given pair, we compute the fully marginalized likelihood (FML) for the hypotheses (1) and (2),
$\mathcal{L}_1$ and $\mathcal{L}_2$.
We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as the scalar quantity to select
candidate comoving pairs, as described in $\sectionname~\ref{sub:selection}$ in
more detail.
To compute these FMLs, the likelihood functions for each star in a pair, $L_i,
L_j$, are marginalized over the (unknown) true distance and 3D velocity for each star in
the pair $(i,j)$.
\begin{align}
  \mathcal{L}_1 &=
    \int \, \dd r_i \, \dd r_j \, \dd^3 \vec{v} \,
    L_i(\vec{\mu_i} \given \vec{v}, r_i, s^2) \,
    L_j(\vec{\mu_j} \given \vec{v}, r_j, s^2) \,
    p(\vec{v}) \, p(r_i \given \varpi_i) \, p(r_j \given \varpi_j) \label{eq:hyp1}\\
  \mathcal{L}_2 &=
    \int \, \dd r_i \, \dd r_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
    L_i(\vec{\mu_i} \given \vec{v}_i, r_i, s^2) \,
    L_j(\vec{\mu_j} \given \vec{v}_j, r_j, s^2) \,
    p(\vec{v}_i) \, p(\vec{v}_j) \, p(r_i \given \varpi_i) \, p(r_j \given \varpi_j). \label{eq:hyp2}
\end{align}
Here, $p(r_i \given \varpi_i)$ is the posterior distribution of distance given parallax measurement
$\varpi_i$ and its Gaussian error $\sigma_{\varpi,i}$.
Note that the FML for the hypothesis (1) involves integration over one velocity $\vec{v}$ that
generates the likelihoods for both stars, $L_i$ and $L_j$.
The marginalization integral for hypothesis (2) can be split into the product of
two simpler integrals $\mathcal{L}_2 = Q(\vec{\mu_i}, \varpi_i) \, Q(\vec{\mu_j}, \varpi_j)$ where
\begin{equation}
  Q(\vec{\mu}, \varpi) = \int \, \dd r \, \dd^3 \vec{v} \, L(\vec{\mu} \given \vec{v}, r, s^2) \,
  p(\vec{v}) \, p(r \given \varpi)
\end{equation}

@@

This section is divided into three parts. First, we discuss and justify a cut of
the likelihood ratio to select candidate comoving pairs.
Second, we present the statistics and properties of our candidate comoving
pairs.
Finally, we describe our catalog of candidate comoving pairs, the main product
of this study.

\subsection{Selecting candidate comoving pairs}
\label{sub:selection}

In this section, we examine the distribution of (log-)likelihood ratios
$\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
for this quantity to select comoving pairs from the initial sample.

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{likelihoodratios.pdf}
  \end{center}
  \caption{%
    Justification of the likelihood ratio cut.
    On the top row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
    changes (left) in slices of \absdvtan\ (sequentially color-coded in pink on the right panel).
    On the bottom row, we show how the distribution of pairs on \absdvtan\ vs. separation changes (right)
    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue on the left).
    The black dots in the upper right panel and
    the black line in the lower left panel correspond to the entire initial sample of pairs.
    For comparison, we also present the likelihood ratio distribution of random pairs
    of stars in the \tgas\ with parallax $S/N>8$ in gray filled histogram.
    Based on this, we choose $\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as our high-confidence
    candidate comoving pairs.
    \label{fig:likelihoodratios}}
    \vspace{1em}
\end{figure*}

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
$\sim 271$k pairs in the initial sample.
As discussed in \sectionname~\ref{sec:data}, we expect a correlation between the likelihood
ratios of pairs, and their distribution on \absdvtan\ vs separation plane.
Specifically, as we sweep through from small \absdvtan\ to large, we expect
the population of pairs to change from genuinely comoving to random.
This becomes clear when we look at the distribution of the likelihood ratios of pairs
in slices of \absdvtan\ (top row of Figure~\ref{fig:likelihoodratios}).
Pairs with $\absdvtan \in (0,2.5)~\kms$ are most likely actual comoving pairs, and
their likelihood ratio distribution is narrowly peaked at $\gtrsim 5$
(darkest pink histogram in the upper left panel of Figure~\ref{fig:likelihoodratios}).
The distribution peaks at lower values and gets broader as \absdvtan\ increases,
and the number of random pairs increasingly dominate.
On the bottom row of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on \absdvtan\ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \sectionname~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as the initial sample ($S/N > 8$).
Shown as the gray filled histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable comoving pairs.

Based on these comparisons, we select candidate comoving pairs with
$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$.
Out of 271,232 pairs in the initial sample, 13,058 pairs (4.8\%)
satisfy this condition.

\subsection{Statistics and properties of the identified comoving pairs}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{graphviz_examples.pdf}
  \end{center}
  \caption{%
    Visualizations of a few example connected components of comoving pairs of stars.
    Each star (node) is marked as a red circle, and a line (edge) is drawn
    between two stars if they are comoving by our selection criteria
    (see \sectionname~\ref{sub:selection}). On left, we show the largest network
    found in this study corresponding to the Pleiades star cluster.
    On right, we show four examples of connected components with varying sizes.
    The connected component on the upper left panel with a size of 58 corresponds
    to NGC~2632, also known as the Beehive cluster.
    \label{fig:graphviz_examples}}
\end{figure*}

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{glon_d_pie.pdf}
  \end{center}
  \caption{%
    Panoramic view of comoving pairs of stars around the Sun. This is a cylindrical
    projection of onto the Galactic plane with the Sun at the origin.
    Pairs in connected components of sizes less than 5 are connected by gray lines.
    Pair in connected components of size $\geq 5$ are plotted with a unique random color,
    and their nodes (stars) are highlighted with small black circles.
    We also show the positions of known Milky Way star clusters
    from \citet{Kharchenko:2016aa} as light blue circles of $10$~pc radius,
    and stars in OB associations from \citet{de-Zeeuw:1999aa} as
    blue crosses.
    Some of the larger connected components are clearly associated with known
    clusters, but we also discover quite a few new comoving groups.
    \label{fig:glon_d_pairlines}}
\end{figure*}

Once we have identified candidate comoving pairs from the initial
sample, these pairs form an undirected graph where stars are nodes, and edges
between the nodes exist for comoving pairs of stars.
A star may have multiple comoving neighbors, and two stars may either
be directly or indirectly connected by a sequence of edges (``path'').
We divide the graph into connected components\footnote{A connected component of an undirected graph $G$
is a subgraph of $G$ in which any two nodes are connected to each other by a path.},
and show the distribution of their sizes in Figure~\ref{fig:hist_ccsize}.
The most common are connected components of size 2, which mean mutually exclusive comoving
pairs. However, it is clear that there are many aggregates of comoving stars
discovered by looking for comoving pairs.
These aggregates are likely moving groups, OB associations, or star clusters.
In 13,058 comoving pairs that we identified,
there are 4,555 connected components among 10,606 unique stars.
The maximum size of the connected components is 151.
We show this largest connected component along with four other
examples of varying sizes in Figure~\ref{fig:graphviz_examples}.
The largest connected component corresponds to the Pleiades open cluster (left panel of Figure~\ref{fig:graphviz_examples})
while the upper left panel of the right column of Figure~\ref{fig:graphviz_examples}
is NGC~2632, another known Milky Way open cluster \citep{Kharchenko:2016aa}.
